Passenger Prediction in Shared Accounts for Flight Ticket Recommendation

### 2016-06-28 Paper.tex
Add files via upload

%
% --- Author Metadata here ---
\conferenceinfo{WOODSTOCK}{'97 El Paso, Texas USA}
%\CopyrightYear{2007} % Allows default copyright year (20XX) to be over-ridden - IF NEED BE.
%\crdata{0-12345-67-8/90/01}  % Allows default copyright data (0-89791-88-6/97/05) to be over-ridden - IF NEED BE.
% --- End of Author Metadata ---

\title{A Context-Aware Method for User Prediction in Shared Accounts for Recommendation}

@@

\maketitle
\begin{abstract}
Traditional recommender systems including collaborative filter and content-based method usually assume every account as a single user. However, under many circumstances, multiple users share a common account. Such as an online shopping account for family members and an online ticket booking account for family or companions. For the former case, service provider can't access user's personal information precisely. For the latter case, user's personal information can be acquired only after the user has submitted the order or checked in. If we can predict current user before he making a choice, then we may provide a more pertinent recommendation. For scenarios where user's within an account can't be identified precisely, there are some methods such as context aware methods and clustering. In this paper, we deal with the situation that user's personal information can be acquired. we propose a probabilistic method for user identification in a shared account with history behaviors and contextual knowledge. We then experimentally demonstrate its effectiveness on a dataset of Ctrip.com. The result show that our method outperforms conventional methods.
\end{abstract}

@@

Recent years, with the repaid development of online travel agency, there are more and more passengers booking tickets or hotels through online service provider. It is quite common that a user books tickets for the whole family or companions. Typically, an user may input his location, destination and departure time, then he will get a list of candidate items. In general, there are dozens of items within a list , all of them are isomorphic. Thus, it takes quite long time for an user to compare and choose the most approximate one. Therefore, a recommender system is quite necessary to be applied for better user experience. Figure 1 shows a list for all candidate flights departure from Beijing to Hong Kong. There are total 156 matching flights in result. The website only provides simple sorting strategies such as departure time, arrival time, price. Thus, it may take quite long time for users to find the appropriate one.\par

@@

Essentially, an item recommender system is used for generating a personalized ranking on a set of items\cite{lv:rec}. Typical recommender systems assume that every account represents a single preference model. However, in many cases, multiple users will share one account. Depending on the service, some service providers can acquire user's personal information within an account precisely while others can't. For most online travel service providers, user's personal information can be acquired, for example, online airline ticket providers require users going aboard to fill in their personal information, and online hotel booking service providers can acquire users' personal information when they check in. It is likely that each user has different preference from each other. If we can identify users within a single account for current contextual environment, we may get the chance to improve accuracy of recommendation.\par
To the best of our knowledge, all existing methods focus on extracting implicit users within an account. Some researches\cite{kabbur:nlmf}\cite{jason:embedding} extract users applying nonlinear latent factorization. Some researches\cite{koen:top-n}\cite{yutaka:modeling} focus on user extraction with no context information. These methods always treat an account as a fixed count of implicit users, and the experiment result can be only evaluated on artificial composite datasets. In this article, we deal with the challenge that service providers can acquire a user's personal information only after he has submitted the order. Our goal is to predict distribution of users in current session based on history behavior and contextual knowledge, thus we can make more pertinent recommendation. To achieve this target, we propose a probabilistic author topic model to analyze users' behavior. We regard all orders in an account as corpus, all users in account as authors, orders as documents, content attributes of the items and contextual knowledge as item list. Author is a multinomial distribution over topics and topic is a multinomial distribution over items. Note that if an order has more than one user, these users are co-authors of this document.\par
In experiment section, we apply our method in a flight ticket order dataset of Ctrip, largest OTA service provider in China. In this dataset, we have submitted order as well as trace log in that session. And the result shows that our method has higher individual identification accuracy and a better recommendation accuracy.\par
This article provides following contributions:\\

@@

\item We provide a probabilistic user identification method for the situation, where personal information within an account can be acquired.
\item We evaluate the accuracy of predicting distribution of user with certain contextual knowledge on a flight ticket order dataset.
\item We apply the predicting result in recommendation, and the result shows that our approach can effectively improve the accuracy.

@@

\textbf{Paper organization.} The rest of paper is structured as follows. We introduce related work in section 2. In section 3, we define the problem and describe the model, besides we show part of process on parameter inference. In section 4, we present dataset, evaluating metric and the result of predicting and recommendation. Finally, we conclude the paper in section 5.

\section{Related Work}
\textbf{Content based recommendation.} In online travel agency fields such as airline tickets or hotel booking, the items are usually highly structured, however, the content of items are quite mutable. Thus we can't propose a general metric to define items. Content-based recommender systems \cite{tech:cset}\cite{lops:handbook}are appropriate for highly structured items. The method needs proper measurements for representing the attributes of items and the profile of user interests, then it matches up the profile of user's interest with attributes of items. In general, a content-based recommender systems consists of three components. Content analyzer extract the structured attributes of items through a series of pre-processing steps. It produces the input for the system, which always needs domain related knowledge. The profile learner analyzes and constructs user's profile of interests by collected user's history behaviors. In addition, this component also relies on feedbacks of users. The filtering component evaluate the relevance of items for particular user through matching them with user's profile. The output for an item can be either binary or continuous, in this stage, some similarity measurements including cosine or euclidean distance can usually be applied.\par
\textbf{Implicit feedback.} Learning a profile of user interests is a form of classification. The training data is labeled user history behavior log. Typically, for explicit feedback, training data can be divided into binary classes, indicating whether the user likes the item or not. The explicit feedback can be acquired by ratings and text comments. However, explicit feedback increases the cognitive load on user, so users may not willing to provide feedback, and it also suffers from the transformation of user interests. In several application domains, a single numeric rating may not be adequate for describing users feeling over an item. Implicit feedback methods\cite{song:mining} are based on assigning a relevance score to specific actions on an item, such as clicking ,collecting, bookmarking, etc. These actions do not require user's involvement directly, so it also has the advantage of being free from the problem of lacking feedback data. Nevertheless, implicit feedback methods suffer from lacking of negative feedbacks and data noises. Based on the weakness of implicit feedback recommender systems, some methods are proposed, such as one class collaborative filtering and content based recommender.\par
\textbf{Probabilistic topic models.} Some information retrieval approaches are involved in  model based recommender systems. Topic models\cite{blei:lda}\cite{mic:atm}\cite{mark:prob} are proposed for automated extraction of useful semantic information from corpus data, which are widely used in information retrieval, document annotation, etc. The main process of models are extracting latent factors from corpus, named topics, which are commonly probabilistic distribution over words. Topic models have three major advantages over other document modeling methods. At first, topic models provide a completely unsupervised approach to extracting topics, thus requiring no document labels and no initialization. Unsupervised methods are necessary for modeling large document corpus, besides, in some fields, with the rapid change of content's topics, predefined topic categories may not reflect the development of specific field. Secondly, latent topics are interpretable respectively, so that extracted results are understandable for users. Finally, each document may consist of multiple topics, document can be considered as probabilistic distribution over topics. Many existing models regard documents as mixtures of topics, such as pLSI, LDA. The author topic model is a generative model that extends LDA to include authorship information for document modeling.\par
\textbf{Shared account recommendation.} Some work aims at addressing the challenge of multiple user sharing a single account. A top-N recommendation for shared accounts approach was proposed by Koen et al.\cite{koen:top-n} It is an item-based top-N collaborative filtering recommender system on binary, positive-only feedback. Another method proposed by Santosh et al.\cite{kabbur:nlmf} models users within an account with a much richer representation. It uses a nonlinear matrix factorization methods for predicting the recommendation score. Yutaka et al.\cite{yutaka:modeling} introduced an approach for modeling multiple users' purchase in a single account using an extended pLSA model. All above methods are applied in scenario that users' personal information can't be acquired precisely.

\section{User Prediction Model}
In this section, we propose a generic model for predict the probabilistic distribution of users within an account and produce the prediction result under certain contextual environment. Finally we integrate user prediction process into recommendation to get a better recommendation. Figure 2 shows an overview of the recommendation process. At first,a dataset for account orders are stored, each account may contains several users, whose identification can be acquired. We use this data to train user prediction model as well as extract preference at user granularity. Before a recommendation process starts, we calculate user distribution probability applying trained model and contextual knowledge of current session. At the final step, we generate a ranked candidate list as the recommendation result based on user prediction result and user's preference.\par
\begin{figure}[!hbt]
\centering
\includegraphics[scale=.40]{overview.eps}
\caption{Overview of user prediction integrated recommendation}
\label{fig:Prob}
\end{figure}\par

\subsection{Model description and notation}
Topic models are widely used in recommender systems. For example, in pLSA model\cite{tomas:coll}, a topic is a multinomial distribution over items, it represents the latent feature. And an account is a multinomial distribution over topics, it represents the preference of an account. For every purchase, we sample a topic z from account a, and take an item from topic z. Model-based recommender algorithms can usually achieve a higher prediction accuracy and reduce dimension of the dataset, in the meanwhile, it has more generality and flexibility.\par
In flight ticket booking field, the concept of item is quite different. Conventional items only have static features, means that the content of an item is less likely to change in the future. For flight tickets, however, the price changes frequently even for the same flight number and class. It is well known that the price factor plays a significant role in users' choice, so we can't consider the ticket of different price as one item. Besides, the amount of tickets for a flight is very limited, which leads to a very sparse user-item matrix, so we can't use a collaborative-filtering like method.\par
Fortunately, the decisive factors of a flight ticket are quite fixed, such as airline, takeoff time, price and class etc. Some factors are discrete while others are continuous. For continuous factors, we try to partition them into discrete non-overlapping intervals. Thus, the number of discrete alternatives of each factor is also very limited. We take orders within each account as a corpus. Our notations are summarized in table 1.\par
\begin{table}[!htbp]
\centering
\caption{Notation}
\begin{tabular}{|c|c|} \hline
$M$ & number of accounts within dataset\\ \hline
$V$ & number of words in vocabulary\\ \hline
$O$ & orders in an account\\ \hline
$P$ & passengers in an account\\ \hline
$P_i$ & passengers for order $O_i$ \\ \hline
$F$ & selected Factors\\ \hline
$V$ & size of vocabulary\\ \hline
$K$ & number of topics\\ \hline
\end{tabular}
\end{table}

In author topic model\cite{deiv:corpora}, a predefined vocabulary is generated containing discrete alternatives and intervals of all selected factors, we use integers to donate each entry in vocabulary. The model discovers both topics distributed in an order and passengers associated with distribution of topics. We treat every order as a bag of words, thus deduces each order to a vector of word counts. An order can be considered as a document with fixed word length, which can be represented as a vector of integers. All passengers corresponding to an order are co-authors for that document. Each passenger is associated with a multinomial distribution over topics and each topic is associated with a multinomial distribution over entries. The model essentially reduces dimension of documents, depending on the amount of topics.\par
For an account M, we generate the observed passenger list $P$. We denote authors' distribution over topics by a $|P| \times K$ matrix $\Theta$. The multinomial distribution can be generated from Dirichlet prior distribution with hyper-parameter $\alpha$. Topics' distribution over words is denoted by a $K \times \sum_{i=1}^{|F|}|F_i| $  matrix $\Phi$, also, distribution can be generated with hyper-parameter $\beta$. Generally, these hyper-parameters needn't be estimated, here, we fix $\alpha$ and $\beta$ at 50/K and 0.01 respectively. \par
To generate a word, we need draw two latent variables, respectively a passenger and a topic. First we draw a passenger uniformly from $P$, then a topic $Z$ from $\Theta_A$ and a word w from $\Phi_Z$. The following process describes generative model mathematically.\par

\begin{enumerate}
\item For each passenger $p=1,\dots,|P|$ draw $\Theta_p \sim Dirichlet(\alpha)$
\item For each topic $t=1,\dots,K$ draw $\Phi_t \sim Dirichlet(\beta)$
\item For each order $o=1,\dots,O$
       \begin{enumerate}[fullwidth,itemindent=1em,label=(\alph*)]
       \item given passengers $P$
       \item For each word $i=1,\dots, N_o$
              \begin{enumerate}[fullwidth,itemindent=2em,label=(\roman*)]
              \item draw a passenger $X_{oi} \sim Uniform(P)$
              \item draw a topic $Z_{oi} \sim Discrete(\theta_{X_{oi}})$
              \item draw a word $w_{oi} \sim Discrete(\phi_{Z_{oi}})$
              \end{enumerate}
       \end{enumerate}
\end{enumerate}
\begin{figure}[!hbt]
\centering
\includegraphics[scale=.55]{f1.eps}
\caption{Probability Graph for author topic model}
\label{fig:Prob}
\end{figure}\par
Figure 3 shows the graphical model corresponding to this process, where shaded and non-shaded nodes represent observed and unobserved variables respectively. It's a plate notation, the number at the bottom of each box indicates repeated time of operation. Arrows represents the conditional dependency between variables. In author topic model, observed variables includes words in an order and the passengers of the order.

\subsection{Parameter inference}
As mentioned above, the author topic model includes two sets of unknown parameters, the $P$ passenger-topic distributions $\theta$, and the $K$ topic-word distributions $\phi$. There are a variety of algorithms that can be applied to estimate the parameters of author topic models. The Expectation-Maximization algorithm and Gibbs sampling\cite{gregor:esti}. Generic EM algorithms are likely to face local maxima and computational inefficient. In this paper, we utilize Gibbs sampling, it does not explicitly estimate parameters, instead, it evaluates posteriori distribution just on drawn passenger $X$ and topic $Z$. Thus it is simple for Dirichlet priors.\par
We can obtain the probability of every word $\mathbf{w}_o$ generated in each order, conditioned on $\Theta$ and $\Phi$ is:\\
\begin{flalign}
\begin{split}
& P(\mathbf{w}_o | \Theta,\Phi,P) =\prod_{i=1}^{N_o}P(w_{oi}|\Theta,\Phi,\mathbf{p}_o) \\
& =\prod_{i=1}^{N_o}\sum_{p=1}^{|P|}\sum_{t=1}^{K}P(w_{oi}|z_{oi}=t,\Phi)
P(z_{oi}=t|x_{oi}=p,\Theta)P(x_{oi}=p|\mathbf{p}_o)\\
& =\prod_{i=1}^{N_o}\frac{1}{|P|}\sum_{p \in p_o}\sum_{t=1}^{K}\phi_{w_{oi}t}\theta_{tp}
\end{split} &
\end{flalign}

@@

\subsection{Passengers Prediction}

@@

%TODO:Conclude the algorithm
In conclusion, the passenger prediction task can be partitioned into two steps. For the first step, a set of decisive factors are extracted by domain field knowledge to generate a vocabulary. And the parameter $\Theta$ and $\Phi$ are trained through several batches of Gibbs sampling. In next step, passengers of an anonymous order can be predicted applying Equation 5. Algorithm 1 describes the predicting process mathematically.\par

@@

\section{Experiments}
In the following subsections, we introduce datasets, baseline method, evaluation metric. Then we evaluate the performance of user prediction and prediction integrated recommendation respectively.
\subsection{Dataset and baseline recommendation method}
In our experiments, we use a flight ticket dataset provided by Ctrip.com, consisting of all submitted orders for duration of two years. An order contains account id and all passengers' id. We use desensitized passenger id to distinguish individuals from each other. Besides, an order contains some travel information that can be considered as decisive factors, such as departure airport, arrival airport, order submitted time, takeoff time, etc. In our experiment, we sample several active users whose amount of history orders reaches a threshold, thus user's preference can be extracted with more confidence. In addition, many researches\cite{amy:guess}\cite{yutaka:modeling} on shared account recommendation generate an artificial dataset by composite single account. We also generate an artificial dataset combining single passengers from different accounts. We extract two passengers from arbitrary accounts and combine them as one account. If these passengers happen to take the same flight, we delete one order randomly. Overviews of two datasets and the meta data of orders are summarized in table 2 and table 3.
\begin{table}[!htbp]
\centering
\caption{Datasets}
\begin{tabular}{|c|c|c|c|} \hline
& \# accounts & \# passengers & \# orders\\ \hline
Real & 4632&7034&38907 \\ \hline
Artificial &1604&3208&29759\\ \hline
\end{tabular}
\end{table}

@@

The meta data of an order can be classified as travel information, context information, individual information and order information. Travel information contains the departure and arrival city, takeoff time. Context information contains time of browsing, other searching and filtering actions. Individual information contains account id, personal information. Order information consists of other content of an order.\par
%TODO: More concise method

With factors mentioned above, we propose a baseline content based recommendation method. First, select decisive factors and partition each factor as discrete variables or intervals, thus we can represent a candidate flight ticket as a vector. Then extract a user's preference from his history orders, for every factor, we generate a vector representing all alternatives and assign each dimension with the percentage of occurrence frequency. For every item in candidate ticket list, we compare it with users' preference model and get a score for every factor based on the frequency percentage. We can sum up score of every factor to get a total score for item, and rank all candidates in the list according to the total score. We can then provide a top-N recommendation.\par

>>>>

%
% --- Author Metadata here ---
%\conferenceinfo{WOODSTOCK}{'97 El Paso, Texas USA}
%\CopyrightYear{2007} % Allows default copyright year (20XX) to be over-ridden - IF NEED BE.
%\crdata{0-12345-67-8/90/01}  % Allows default copyright data (0-89791-88-6/97/05) to be over-ridden - IF NEED BE.
% --- End of Author Metadata ---

\title{Passenger Prediction in Shared Accounts for Flight Ticket Recommendation}

@@

\maketitle
\begin{abstract}
Personalized recommendation is needed for online flight ticket booking service because it is a difficult task for a traveller to select the flight when the number of available flights is large. Traditionally, we can recommend flights for a user based on his historical orders from his account. However, people often book tickets for companions through his account. In this case, the preferences of other travellers should also be considered. Unfortuanately, before placing the order, people will not provider passengers' information. Therefore, we propose a probabilistic method for passenger prediction based on historical behaviors and contextual knowledge. We then experimentally demonstrate its effectiveness on a real dataset. The result shows that our method outperforms conventional methods.
\end{abstract}

@@

Recent years, with the rapid development of online travel agency, there are more and more passengers booking flight tickets through online service providers. Typically, an user may input his location, destination and departure time, then he will get a list of candidate flights. In general, there are dozens of items within a list, all of them are isomorphic. Thus, it takes quite long time for the user to compare and choose the most appropriate one. Therefore, a recommender system is quite necessary to be applied for better user experience. Figure 1 shows a list for all candidate flights departure from Beijing to Hong Kong when input a query. There are total 156 matchable flights in result. The website only provides simple sorting strategies such as departure time, arrival time, price. Thus, it may take quite long time for users to find the appropriate one.\par

@@

Essentially, an item recommender system is used for generating a personalized ranking on a set of items\cite{lv:rec}. Typical recommender systems assume that every account represents a single preference model. However, in online flight ticket booking, it is quite common that a user books tickets for the whole family or companions.  Obviously, every passenger has his own preferences. Unfortuanately, before placing the order, people will not provider passengers' information.  Therefore, if we want to improve the accuracy of recommendation, we have to predict who are going to fly.  \par
To predict who is using the account to place orders is regarded as shared account problem. Some approaches have been proposed to extract implicit user information within an account. For example,  in \cite{kabbur:nlmf}\cite{jason:embedding} nonlinear latent factorization is applied to  extract user information. However, these methods only rely on the analysis of relationships between products and users \cite{koen:top-n}\cite{yutaka:modeling}, and context information that can help differentiate users is neglected. Besides, current methods always treat an account as a fixed number of implicit users, and the experiment result can be only evaluated on artificial composite datasets. 
The flight booking service providers can acquire user's personal information only after he has submitted the order. Thus, our goal is to predict distribution of passengers in current session based on historical behavior and contextual knowledge so that we can make more pertinent recommendations. To achieve this target, we adopt the probabilistic author topic model to analyze passengers' behaviors. We regard all orders in an account as corpus, all passengers appearing in the account as authors, orders as documents, content attributes of the items and contextual knowledge as item list. Note that if an order has more than one passenger, they are co-authors of this document.\par
In experiment section, we apply our method to a real flight ticket order dataset.  And the result shows that our method has higher identification accuracy and a better recommendation accuracy.\par
This paper has following contributions:\\

@@

\item We provide a probabilistic passenger identification method for online flight booking service.
\item We apply the passenger prediction results to flight recommendation.
\item We evaluate our approach on a real flight order dataset, and the result shows that our approach can effectively improve the recommendation accuracy.

@@

\textbf{Paper organization.} The rest of paper is structured as follows. We introduce related work in Section 2. In Section 3, we define the problem and describe the model. Besides, we also discuss how to do parameter inference. In section 4, we introduce the experiment dataset, evaluation metrics and the experimental results of passenge prediction and flight recommendation. Finally, we conclude the paper in section 5.

\section{Related Work}
\textbf{Content based recommendation.} In online travel service industry such as airline tickets or hotel booking, the item information is usually highly structured, however, this information is quite mutable. Content-based recommender systems \cite{tech:cset}\cite{lops:handbook}are appropriate for highly structured items. The method needs proper representations of the attributes of items and the profile of user interests, then it tries to match  the profile of user's interests with attributes of items. In general, a content-based recommender system consists of three components. A content analyzer extracts structured attributes of items through a series of pre-processing steps. It produces the inputs for the system, which always need domain related knowledge. A profile learner analyzes and constructs user's profile of interests by collecting user's historical behaviors. In addition, this component also relies on feedbacks of users. A filtering component evaluates the relevance of items to particular users by matching them with users' profiles. The output for an item can be either binary or continuous, in this stage, some similarity measurements including cosine or euclidean distance can usually be applied.\par

\textbf{Flight Ticket Recommendation.} There is little work focusing on personalized flight ticket recommendation. Many online flight ticket booking agencies provide sorting, filtering strategies to improve user experience, such as sorting by price or takeoff time and filtering by airline or departure airport. Others also provide additional information that can help users, for example, placing the hottest flights at top or highlighting some of them. These methods do make sense in helping users find appropriate flights within less time. However, they can not provide personlized recommendations. In 2004, Lorcan Coyle\cite{lor:flight} proposed a Personal Travel Assistant(PTA). PTA is a method of case based reasoning, which takes user's historical orders as cases and recommends the most similar candidate flight to user's query. In addition, PTA also proposes attribute weight training and case storage method.\par
\textbf{Implicit feedback.} The user's feedback can be divided into implict  or explict type. The explicit feedback can be acquired from users' ratings or text comments. However,  users may not willing to provide explict feedback since it needs additional efforts. Moreover, in many applications, a single numeric rating may not be adequate for describing users feeling to an item. Implicit feedback methods\cite{song:mining} try to infer users' attitudes to an item based on his behaviors, such as clicking ,collecting or bookmarking et al. It has the advantage of being free from the problem of lacking feedback data. Nevertheless, implicit feedback based recommender suffers from problems of lack of negative feedbacks and data noises. In order to provide recommendations when only implict feedbacks are available, some methods are proposed, such as one class collaborative filtering and content based recommender.\par
\textbf{Probabilistic topic models.}  Topic models\cite{blei:lda}\cite{mic:atm}\cite{mark:prob} are proposed for automated extraction of useful semantic information from corpus data, which are widely used in information retrieval, document annotation, etc. The main step is extracting latent factors from corpus, named topics, which are commonly probabilistic distribution over words. Topic models have three major advantages over other document modeling methods. At first, topic models provide a completely unsupervised approach to extract topics, thus requiring no document labels and no initialization. Unsupervised methods are necessary for modeling large document corpus, besides, in some fields, with the rapid change of content's topics, predefined topic categories may not reflect the development of specific field. Secondly, latent topics are interpretable respectively, so that extracted results are understandable for users. Finally, each document may consist of multiple topics, document can be considered as probabilistic distribution over topics. Many existing models regard documents as mixtures of topics, such as pLSI, LDA. The author topic model is a generative model that extends LDA to include authorship information for document modeling. In our approach, topic model is innovatively applied to indentify users. \par
\textbf{Shared account recommendation.} Some work aims at addressing the challenge of identifying current users who share a single account. A top-N recommendation for shared accounts approach was proposed by Koen et al.\cite{koen:top-n} It is an item-based top-N collaborative filtering recommender system on binary, positive-only feedback. Another method proposed by Santosh et al.\cite{kabbur:nlmf} models users within an account with a much richer representation. It uses a nonlinear matrix factorization methods for predicting the recommendation score. Yutaka et al.\cite{yutaka:modeling} introduced an approach for modeling multiple users' purchase in a single account using an extended pLSA model. All above methods are applied in scenario that users' personal information can't be acquired explicitly.

\section{Passenger Prediction Model}
In this Section, we propose a generic model for predicting the probabilistic distribution of passengers under certain contextual environment. Finally we integrate user prediction into recommendation to get higher accuracy. Figure 2 shows an overview of the recommendation process. At first,a dataset for orders are stored, each account may contains several passengers. We use this data to train user prediction model as well as extract preference at user granularity. Before a recommendation process starts, we calculate user distribution probability applying trained model and contextual knowledge of current session. At the final step, we generate a ranked candidate list as the recommendation result based on user prediction result and user's preference.\par
\begin{figure}[!hbt]
\centering
\includegraphics[scale=.40]{overview.eps}
\caption{Overview of flight recommendation based on passenger prediction}
\label{fig:Prob}
\end{figure}\par

\subsection{Model description and notation}
Topic models are widely used in recommender systems. For example, in pLSA model\cite{tomas:coll}, a topic is a multinomial distribution over items, it represents the latent feature. And an account is a multinomial distribution over topics, it can represent the preference of an account. Each purchase can be regared a sample that selects a topic z from account a, and takes an item from topic z.\par

When topic model is applied to flight recommendation, there are some challenges. Firstly, the concepts behind flight tickets are quite different to be modeled. Conventional items only have static features, means that the content of an item is less likely to change in the future. For flight tickets, however, the price changes frequently even for the same flight number and class. Since the price factor plays a significant role in users' choice, so we can't consider the ticket of different price as one item. Secondly, the amount of tickets for a flight is limited, which leads to a very sparse user-item matrix so that we can't use a collaborative-filtering like method directly.\par

Fortunately, the decisive factors of a flight ticket are quite fixed, such as airline, takeoff time, price and class etc. Some factors are discrete while others are continuous. For continuous factors, we can partition them into discrete non-overlapping intervals applying . Thus, the number of discrete alternatives of each factor is also very limited. We take orders within each account as a corpus. Our notations are summarized in table 1.\par
\begin{table}[!htbp]
\centering
\caption{Notation}
\begin{tabular}{|c|c|} \hline
$M$ & number of accounts within dataset\\ \hline
$V$ & number of words in vocabulary\\ \hline
$O$ & orders in an account\\ \hline
$P$ & passengers in an account\\ \hline
$P_i$ & passengers for order $O_i$ \\ \hline
$F$ & selected Factors\\ \hline
$V$ & size of vocabulary\\ \hline
$K$ & number of topics\\ \hline
\end{tabular}
\end{table}

In author-topic model\cite{deiv:corpora}, a predefined vocabulary is generated containing discrete alternatives and intervals of all selected factors, we use integers to donate each entry in vocabulary. The model discovers both topics distributed in an order and passengers associated with distribution of topics. We treat every order as a bag of words, thus deduces each order to a vector of word counts. An order can be considered as a document with fixed word length, which can be represented as a vector of integers. All passengers corresponding to an order are co-authors for that document. Each passenger is associated with a multinomial distribution over topics and each topic is associated with a multinomial distribution over entries. The model essentially reduces dimensions of documents, depending on the amount of topics.\par

For an account M, we generate the observed passenger list $P$. We denote authors' distribution over topics by a $|P| \times K$ matrix $\Theta$. The multinomial distribution can be generated from Dirichlet prior distribution with hyper-parameter $\alpha$. Topics' distribution over words is denoted by a $K \times \sum_{i=1}^{|F|}|F_i| $  matrix $\Phi$, also, distribution can be generated with hyper-parameter $\beta$. Generally, these hyper-parameters needn't be estimated, here, we fix $\alpha$ and $\beta$ at 50/K and 0.01 respectively. \par

To generate a word, we need draw two latent variables, respectively a passenger and a topic. First we draw a passenger uniformly from $P$, then a topic $Z$ from $\Theta_A$ and a word w from $\Phi_Z$. The following process describes generative model mathematically.\par

\begin{enumerate}
\item For each passenger $p=1,\dots,|P|$ draw $\Theta_p \sim Dirichlet(\alpha)$
\item For each topic $t=1,\dots,K$ draw $\Phi_t \sim Dirichlet(\beta)$
\item For each order $o=1,\dots,O$
       \begin{enumerate}[fullwidth,itemindent=1em,label=(\alph*)]
       \item given passengers $P$
       \item For each word $i=1,\dots, N_o$
              \begin{enumerate}[fullwidth,itemindent=2em,label=(\roman*)]
              \item draw a passenger $X_{oi} \sim Uniform(P)$
              \item draw a topic $Z_{oi} \sim Discrete(\theta_{X_{oi}})$
              \item draw a word $w_{oi} \sim Discrete(\phi_{Z_{oi}})$
              \end{enumerate}
       \end{enumerate}
\end{enumerate}
\begin{figure}[!hbt]
\centering
\includegraphics[scale=.55]{f1.eps}
\caption{Probability Graph for the author-topic model}
\label{fig:Prob}
\end{figure}\par
Figure 3 shows the graphical model corresponding to this process, where shaded and non-shaded nodes represent observed and unobserved variables respectively. It's a plate notation, the number at the bottom of each box indicates repeated times of operation. Arrows represents the conditional dependency between variables. In the author-topic model, observed variables include words in an order and the passengers of the order.

\subsection{Parameter inference}
As mentioned above, the author-topic model includes two sets of unknown parameters, the $P$ passenger-topic distributions $\theta$, and the $K$ topic-word distributions $\phi$. There are a variety of algorithms that can be applied to estimate the parameters of author-topic models, the Expectation-Maximization algorithm and Gibbs sampling\cite{gregor:esti}. Generic EM algorithms are likely to face local maxima and computational inefficient. In this paper, we utilize Gibbs sampling, it does not explicitly estimate parameters, instead, it evaluates posteriori distribution just on drawn passenger $X$ and topic $Z$. Thus it is simple for Dirichlet priors.\par
We can obtain the probability of every word $\mathbf{w}_o$ generated in each order, conditioned on $\Theta$ and $\Phi$ is:\\
\begin{flalign}
\begin{split}
& P(\mathbf{w}_o | \Theta,\Phi,P) =\prod_{i=1}^{N_o}P(w_{oi}|\Theta,\Phi,\mathbf{p}_o) \\
& =\prod_{i=1}^{N_o}\sum_{p=1}^{|P|}\sum_{t=1}^{K}P(w_{oi}|z_{oi}=t,\Phi)
P(z_{oi}=t|x_{oi}=p,\Theta)P(x_{oi}=p|\mathbf{p}_o)\\
& =\prod_{i=1}^{N_o}\frac{1}{|P|}\sum_{p \in p_o}\sum_{t=1}^{K}\phi_{w_{oi}t}\theta_{tp}
\end{split} &
\end{flalign}

@@

\subsection{Passenger Prediction}

@@

In conclusion, the passenger prediction task can be partitioned into two steps. For the first step, a set of decisive factors are extracted by domain field knowledge to generate a vocabulary. And the parameter $\Theta$ and $\Phi$ are trained through several batches of Gibbs sampling. In the second step, passengers of an anonymous order can be predicted applying Equation 5. Algorithm 1 describes the predicting process mathematically.\par

@@

\section{Experiments}
In the following subsections, we introduce datasets, preference-based recommendation method and evaluation metrics. Then we evaluate the performance of user prediction and passenger prediction based recommendation respectively.
\subsection{Dataset and preference-based recommendation method}
In our experiments, we use a real flight ticket dataset, consisting of all submitted orders for two years. An order contains account id and all passengers' id. We use desensitized passenger id to distinguish individuals from each other. Besides, an order contains some travel information that can be considered as decisive factors, such as departure airport, arrival airport, order submitted time, takeoff time, etc. In our experiment, we select active users whose amount of history orders reaches a threshold, thus user's preference can be extracted with more confidence. In addition, many researches\cite{amy:guess}\cite{yutaka:modeling} on shared account recommendation generate an artificial dataset by compossing single accounts' data. We also generate an artificial dataset combining single passengers from different accounts. We extract two passengers from arbitrary accounts and combine them as one account. Overviews of two datasets and the meta data of orders are summarized in table 2 and table 3.
\begin{table}[!htbp]
\centering
\caption{Datasets}
\begin{tabular}{|c|c|c|c|} \hline
& \# accounts & \# passengers & \# orders\\ \hline
Real & 4632&7034&38907 \\ \hline
Artificial &1604&3208&29759\\ \hline
\end{tabular}
\end{table}

@@

The meta data of an order includes travel information, context information, individual information and order information. Travel information contains the departure and arrival city, takeoff time. Context information contains time of browsing, searching and filtering actions. Individual information contains account id, personal information. Order information consists of price, seat class, flight number, airline company  et al.\par

With factors mentioned above, we propose a user preference based recommendation method. First, select decisive factors and partition each factor's possible values to different levels, thus we can represent a candidate flight ticket as a vector. Then extract a user's preference from his historical orders, for every factor, we generate a vector representing all alternatives and assign each dimension with the percentage of occurrence frequency. For every item in candidate ticket list, we compare it with users' preference model and get a score for every factor based on the frequency percentage. We can sum up score of every factor to get a total score for item, and rank all candidates in the list according to the total score. We can then provide a top-$N$ recommendation.\par

In addition, different users may cares different factors, thus we introduce a weight vector for every account. The evaluation of this weight vector contains two steps, initializing and training, respectively. At the initializing step, we estimate how much a user focuses on a factor through the information entropy, as mentioned in \cite{yang:pers}. We believe that if a user focuses on a factor, his behavior will be less random. The entropy $H$ is a measure to describe the uncertainty for discrete random variable $X$ and probability mass function $P(X)$. Here is the mathematical representation for entropy.\\
\begin{equation}
H(X)=E[-lnP(X)]= - \sum_{i-1}^n P(x_i)log_bP(X_i)
\end{equation}\par
The value $b$ is the base of logarithm, and is usually 2. The entropy of every factor can be calculated and initialized as the weight value after normalization. Considering that different factors have a variety quantity of alternatives and the user may concern different factors some time later, we update the weight vector once the user submits an order. For each submission, we have already produced a ranked list for recommendations, if the user doesn't choice an item (i.e, a ticket) from top-$N$ items, we consider this recommendation fails, so that the weight vector need be retrained to make the rank of selected item improved. The concept of weight training is mentioned in \cite{lor:flight}. By performing weight training for factors, we hope that the result of recommendation getting better in the future. The training procedure is quite straightforward. At first a learning rate should be set, such as $\alpha = 0.1$ . If the selected item has lower similarity for factor $f$ than items ranked before it, the weight value of this factor should be decreased. Otherwise, the weight value should be increased. To avoid the problem of over-fitting, we define a maximum number of iteration step. So the stop criteria is either $1)$ the rank of selected items can't be improved anymore or $2)$ the procedure reaches the iteration time limitation.\\
The preference based recommendation algorithm is described as Algorithm 3. First extracting a user's preference  $P$ and initializing weighted list $W$ through entropy. Items in candidate list are ranked based on inner product of $W$ and corresponding value of factor. After every recommendation, the list is retrained and stored depending on the item that user actually chosen.

\begin{algorithm}[htb]
\caption{Preference-Based-Recommendation}
\label{alg1}
\begin{algorithmic}[1]
\REQUIRE
User's ID $u$ \par
User's history order list $O$ \par
Candidate item list $C$
\ENSURE Ranked candidate list $R$
\STATE Define a factor list $F$;
\STATE $P \leftarrow extractUserPref(O,F)$;
\STATE $W \leftarrow fetchWeightedList(u)$;
\IF{$W$ is $null$}
\STATE initialize $W$ by entropy;
\ENDIF
\STATE $R \leftarrow \emptyset$;
\FORALL { $c : C$} 
\STATE append $(c,weightedScore(c,W))$ to R;
\ENDFOR 
\STATE sort $R$ by score;
\STATE $trainWeightedList(W,R,P)$;
\STATE $saveWeightedList(u,W)$;
\RETURN $R$;
\end{algorithmic} 
\end{algorithm}
